Planetary perturbers: Flaring star-planet interactions in Kepler and TESS

### 2023-10-12 src/tex/ms.tex
[showyourwork] overleaf sync

% Abstract with filler text
\begin{abstract}
    In many star-planet systems discovered so far, the innermost planet orbits within only a few stellar radii. In these systems, planets \textcolor{red}{could} become in-situ probes of the extended stellar magnetic field. Because they disturb the field as they move, they are expected to trigger flares on the stellar surface. \textcolor{red}{If and how their energy and morphology differs from intrinsic stellar flares is poorly constrained.} However, as we expect planet-induced flares to be associated with the planet's orbital period $P_{\rm orb}$, we can identify their presence from a clustering of flares in phase with the planet's orbital phase. We used the excellent phase coverage from Kepler and the Transiting Exoplanet Survey Satellite to find flaring star-planet systems, compile a catalog of all their flares, and measure how much they cluster in orbital phase.

    In the 1811 searched systems, we found 25 single stars with three or more flares each. We quantified the significance of the clustering in each system, and compared it against the theoretically expected power of magnetic interaction that leads to planet-induced flaring. Most systems do not show any clustering, consistent with low expected power. Those we expect to show planet-induced flaring fall on two branches. An inactive one, without any signs of clustering, and a tentative active one, where the clustering becomes more pronounced as the expected power of interaction increases. The flares in HIP 67522 are prominently clustered ($p<0.006$). This young Hot Jupiter system is the most promising candidate for magnetic star-planet interaction in our sample.

\end{abstract}

@@

\textcolor{red}{ignoring factors of the order of 1.}  $B_{\rm p}$ and $B_*$ are the planetary and stellar field strengths at their poles, respectively; $v_{\rm rel}$ is the relative velocity between the stellar rotation and the planet's orbit at the semi-major axis $a$; and $R_*$ and $R_{\rm p}$ are the stellar and planetary radii, respectively. \textcolor{red}{We choose $n=1$ to quantify the strongest possible dependence of the dissipated power on stellar radius and star-planet distance.} Since the planetary fields are unknown, we ignore them in the calculation. We also do not know the stellar field strength at the pole for the vast majority of the stars, so we adopt \textcolor{red}{an average value of 15 per cent of the} surface field strength~(see Section~\ref{sec:methods:bfield}) as a proxy.

@@

Third, we estimate the power of SPI generated by the Alfv\'en wing mechanism ~\citep{saur2013magnetic,kavanagh2022radio},\textcolor{red}{ combining Eqn.~8 for the power of interaction and Eqn.~11 for the magnetopause radius of the planet from\citet{kavanagh2022radio} to arrive at:}

@@

\textcolor{red}{Here, we assume that the magnetopause radius extends to where the magnetic pressures of planet and stellar magnetic wind -- which is dominated by the magnetic field in de sub-Alf\'enic regime -- are equal. We also replace the stellar wind density $\rho_W$ with its proportionality to $\rho_* (R_* / a)^{2}$, and stellar wind magnetic field at the planetary radius $B_W$ with the surface field strength $B_* (R_* / a)^{3}$, which result in the total $R_*^2 a^{-2}$ factor in the above equation. Since the stellar wind density at the stellar surface $\rho_*$ is poorly constrained empirically~\citep{vidotto2021evolution}, we follow~\citet{alvarado-gomez2020tuning,kavanagh2021planetinduced}, and fix $\rho_*=2\cdot10^{10}\rm{cm}^{-3}$. The angle between the wind magnetic field vector and the velocity vector of the planet $\theta$ is also unknown, so we set $\sin\theta=1$ to obtain an upper limit. In reality, $\theta$ will vary both between systems, and over time in any individual system. Omitting $\rho_*$ assumes that all stellar wind densities are the same in our sample, which is not true in reality. The dependence on $\rho_*$ is stronger than that on $B_*$, but weaker than on every other parameter, so it may dominate the magnetic field dependence.}

@@

Lacking $B_{\rm p}$, and various additional factors that appear in $P_{\rm spi,sb},P_{\rm spi,sb0},P_{\rm spi,aw}$, and  $P_{\rm spi,aw0}$, such as the angle between the relative velocity and magnetic field of the stellar wind~\citep{kavanagh2022radio}, or the force-free parameter~\citep{lanza2012starplanet}, we restrict ourselves to differential estimates, and normalize all $P_{\rm spi}$ to AU Mic.

@@

\textcolor{red}{It is important to note that in \citet{chebly2023numerical} they assume a consistent Poynting flux per unit magnetic field strength for all stars, using the value of the Sun as a reference. The size of the Alfvén surface depends strongly on this assumption. However, any variation in the value between different stars would lead to deviations from the law proposed in \citet{chebly2023numerical}, resulting in corresponding changes in the dimensions of the Alfvén surface.
}

@@

\appendix

>>>>

% Abstract with filler text
\begin{abstract}
    In many star-planet systems discovered so far, the innermost planet orbits within only a few stellar radii. In these systems, planets \textcolor{red}{could} become in-situ probes of the extended stellar magnetic field. Because they disturb the field as they move, they are expected to trigger flares on the stellar surface. \textcolor{red}{If and how their energy and morphology differs from intrinsic stellar flares is poorly constrained.} However, as we expect planet-induced flares to be associated with the planet's orbital period $P_{\rm orb}$, we can identify their presence from a clustering of flares in phase with the planet's orbital phase. We used the excellent phase coverage from Kepler and the Transiting Exoplanet Survey Satellite to find flaring star-planet systems, compile a catalog of all their flares, and measure how much they cluster in orbital phase. In the 1811 searched systems, we found 25 single stars with three or more flares each. We quantified the significance of the clustering in each system, and compared it against the theoretically expected power of magnetic interaction that leads to planet-induced flaring. Most systems do not show any clustering, consistent with low expected power. Those we expect to show planet-induced flaring fall on two branches. An inactive one, without any signs of clustering, and a tentative active one, where the clustering becomes more pronounced as the expected power of interaction increases. The flares in HIP 67522 are prominently clustered ($p<0.006$). This young Hot Jupiter system is the most promising candidate for magnetic star-planet interaction in our sample.

\keywords{stars: flare, planet–star interactions, stars: magnetic field, stars: low-mass, stars: activity}
\end{abstract}

@@

\textcolor{red}{ignoring factors of the order of 1.}  $B_{\rm p}$ and $B_*$ are the planetary and stellar field strengths at their poles, respectively; $v_{\rm rel}$ is the relative velocity between the stellar rotation and the planet's orbit at the semi-major axis $a$; and $R_*$ and $R_{\rm p}$ are the stellar and planetary radii, respectively. \textcolor{red}{We choose $n=XXXX$, which best reproduces observations of XXXX in~\citet{lanza2012starplanet}. Since the planetary fields are unknown, we adopt $B_{\rm p}=1\,$G for simplicity.} We also do not know the stellar field strength at the pole for the vast majority of the stars, so we adopt \textcolor{red}{an average value of 15 per cent of the} surface field strength~(see Section~\ref{sec:methods:bfield}) as a proxy.

@@

Third, we estimate the power of SPI generated by the Alfv\'en wing mechanism ~\citep{saur2013magnetic,kavanagh2022radio}, \textcolor{red}{ combining Eqn.~8 for the power of interaction and Eqn.~11 for the magnetopause radius of the planet from\citet{kavanagh2022radio} to arrive at:}

@@

\textcolor{red}{Here, we assume that the magnetopause radius extends to where the magnetic pressures of planet and stellar magnetic wind are equal, and the wind is dominated by the magnetic field in de sub-Alf\'enic regime. We approximate the stellar wind density $\rho_W$ with a proportionality to $\rho_* (R_* / a)^{2}$, where $\rho_*$ is the stellar wind density at the base of the corona. For the stellar wind magnetic field at the planetary radius $B_W$, we insert $B_* (R_* / a)^{3}$, i.e., the field strength decreasing radially outward from the surface as a dipole. This produces an upper bound on the field strength at the planet's distance because more complex fields would drop off more steeply. The last two replacements result in the total $R_*^2 a^{-2}$ factor in the above equation. Since $\rho_*$ is poorly constrained empirically~\citep{vidotto2021evolution}, we adopt $\rho_*=2\cdot10^{10}\rm{cm}^{-3}$, a value often used for the Sun~(XXXX), and for stellar simulations~\citep{alvarado-gomez2020tuning,kavanagh2021planetinduced}. The angle between the wind magnetic field vector and the velocity vector of the planet $\theta$ is also unknown, so we set $\sin\theta=1$ to obtain an upper limit. In reality, $\theta$ will vary both between systems, and over time in any individual system. Omitting $\rho_*$ assumes that all stellar wind densities are the same in our sample, which is not true in reality. The dependence on $\rho_*$ is stronger than that on $B_*$, but weaker than on every other parameter, so it may dominate the magnetic field dependence.}

@@

% Lacking $B_{\rm p}$, and various additional factors that appear in $P_{\rm spi,sb},P_{\rm spi,sb0},P_{\rm spi,aw}$, and  $P_{\rm spi,aw0}$, such as the angle between the relative velocity and magnetic field of the stellar wind~\citep{kavanagh2022radio}, or the force-free parameter~\citep{lanza2012starplanet}, we restrict ourselves to differential estimates, and normalize all $P_{\rm spi}$ to AU Mic.

@@

\textcolor{red}{It is important to note that \citet{chebly2023numerical} assume a consistent Poynting flux per unit magnetic field strength for all stars, using the value of the Sun as a reference. The size of the Alfv\'en surface depends strongly on this assumption. However, any variation in the value between different stars would lead to deviations from the law proposed in \citet{chebly2023numerical}, resulting in corresponding changes in the dimensions of the Alfvén surface.}

@@

\appendix
\restartappendixnumbering
